Detailed structure of phospholipid bilayers with negatively charged lipids and salt concentrations from molecular dynamics simulations including the effects of electronic polarization

### 2019-05-13 Manuscript/manuscript.tex
Edited section Results: agreement w exp. and added dihedral angle data (associated plot TBA)

\subsection{Structural parameters of a pure ECC-POPS model membrane: Agreement with experiments}

The X-ray scattering form factors and C-H bond order parameters from NMR experiments are good measures to validate
the lipid bilayer structure in MD simulations against experiments because they
are related to the bilayer dimensions (area per molecule and thickness) and
conformational fluctuations of individual lipids, respectively, and 
can be directly calculated from MD simulations~\cite{ollila16}.
The X-ray scattering form factors and C-H bond order parameters in
the headgroup and glycerol backbone region of POPS lipid bilayer are already available in the literature \cite{kucerka14,NMRlipidsIV}.
Here, we measure also the C--H bond order parameters of the acyl chain region using natural abundance $^{13}$C NMR.
We first determined the order parameters of each peak corresponding different C--H bonds detected
in the first dimension (Fig. \ref{R-PDLF}) from the dipolar slices in the indirect dimension (Fig. \ref{R-PDLFslices}).
Because the exact assignment of these peaks was not possible, we assigned the
acyl chain order parameters to mimic the profile predicted by the MD simulations (Fig. \ref{simVSexpNOions_POPS}).
\todo{Tiago has to check and update this and figures in SI.}

@@

The Lipid17/Dang simulation give larger acyl chain order parameters and smaller area per molecule than the
experimental values (Fig. \ref{simVSexpNOions_POPS} and table \ref{tab:apls}),
indicating that this simulation predicts too dense POPS lipid bilayer. This is also in line with
the discrepancy observed in the X-ray form factor between this simulation and experiment (Fig. \ref{simVSexpNOions_POPS}).
In previous study \cite{NMRlipidsIV}, the Lipid17 force field with {\AA}qvist ions~\cite{aqvist90}
gave a larger area per molecule (57 \AA$^2$) for POPS being closer to the experimental value,
which is probably due to the weaker counterion binding affinity of {\AA}qvist ions.
However,  {\AA}qvist ions exhibit artificial aggregation of ions at larger salt concentrations \cite{kohagen16, chen07, NMRlipidsIV}.
Therefore, the Dang ions with more realisitic bulk properties are used here to analyze the effect of ECC to the properties of POPS lipid bilayer. 
The ECC-POPS simulation gives a good agreement with the X-ray scattering form factors and experimental
acyl chain order parameters (see Fig.~\ref{simVSexpNOions_POPS}), indicating the the bilayer dimensions and acyl chain
conformations are well described by the force field. The area per lipid from ECC-POPS simulation is slightly
smaller than the value reported from SDP model \cite{kucerka14} (table \ref{tab:apls}), but the values
are within each others error bars.

% In order to obtain the area per lipid from experiments, modeling is used on top of the scattering factors \citep{kucerka14}. 
%In ocntrast, this property is easily extracted directly from simulations with flat phospholipid bilayers. 
%We compare the values from experiments and simulations in Table~\ref{tab:apls}. 
%While the agreement between the scattering form factors from the simulation of a pure POPS bilayer and experiment 
%are excellent (Fig.~\ref{simVSexpNOions_POPS}), there is a non-negligible difference between the values of the area per lipid in Table~\ref{tab:apls}. 
%Since both values are derived from the scattering form factors through modeling of the electron density of the bilayer,
%we cannot decide, which of the values is more reliable.

\begin{table}[tb!] 
\centering
  \caption{Area per lipid (APL) of POPS bilayers from simulations with different force fields at 298~K with \ce{Na^+} counterions. \label{tab:apls} } 
  \begin{tabular}{l|c } 
    \multicolumn{2}{c}{POPS} \\
    model          & APL / Å$^2$    \\ 
    \hline 
    Lipid17/Dang              & 53.5$\pm$ 1.7   \\ 
%    Lipid17/ff99              & 57.9$\pm$ 1.7   \\ 
%    \hline 
    ECC-POPS/ECC-ions         & 59.8$\pm$ 1.8   \\ 
%    \hline 
    experiment (SDP model) \citep{kucerka14} & 62.7$\pm$ 1.3  \\ 
%    \hline 
  \end{tabular} \\
\end{table}

Despite the success in simulating bilayer dimensions and acyl chain conformations,
lipid force fields have typically problems in capturing the correct glycerol backbone and
headgroup order parameters and conformations \cite{botan15,ollila16,NMRlipidsIV}.
This is the case also for Lipid17 POPS simulations, where the headgroup and glycerol backbone
order parameters are quite far from experimental values with both Dang (Fig. \ref{simVSexpNOions_POPS})
and {\AA}qvist \cite{NMRlipidsIV} ions. The values from ECC-POPS simulation are closer,
but not in full agreement with experiments (Fig. \ref{simVSexpNOions_POPS}).
Especially, the experimentally observed more negative values of the $\beta$-carbon
and the larger forking (difference of order parameters between hydrogens attached to the
same carbon \cite{ollila16}) of the $\alpha$-carbon in POPS than in POPC \cite{NMRlipidsIV}
are not reproduced by neither of the models here. These differences are previously related
to a more rigid structure of PS headgroup and the stiffness of the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$
dihedral, which is \todo{the dihedral should be plotted and sentence finished.}

%The head group order parameters $\alpha$ and $\beta$ are highly relevant for this work,
%as they are being used in the electrometer concept \cite{altenbach84, catte16, melcr18}.
%For POPC in pure water, the order parameter $\beta$ agrees well with the experiment, while the order parameter $\alpha$ is somewhat lower.

In conclusion, the ECC-POPS model reproduces the lipid bilayer dimensions and
acyl chain conformations in good agreement with experiments, but has room for
improvement in the glycerol backbone and headgroup region. Therefore,
its structural accuracy is in the same level as in other state-of-the-art
lipid force fields \cite{botan15, ollila16, Pluhackova2016, NMRlipidsIV}.

>>>>

\subsection{ECC-POPS improves agreement in experimental structural parameters of a pure POPS bilayer }

The X-ray scattering form factors and C-H bond order parameters from NMR experiments are good measures to validate
the lipid bilayer structure in MD simulations against experiments because they
are related to the bilayer dimensions (area per molecule and thickness) and
conformational fluctuations of individual lipids, respectively, and 
can be directly calculated from MD simulations~\cite{ollila16}.
The X-ray scattering form factors and C-H bond order parameters in
the headgroup and glycerol backbone region of POPS lipid bilayer are already available in the literature \cite{kucerka14,NMRlipidsIV}.
Here, we measure also the C--H bond order parameters of the acyl chain region using natural abundance $^{13}$C NMR.
We first determined the order parameters of each peak corresponding to different C--H bonds detected
in the first dimension (Fig. \ref{R-PDLF}) from the dipolar slices in the indirect dimension (Fig. \ref{R-PDLFslices}).
Because the exact assignment of these peaks was not possible, we assigned the
acyl chain order parameters to mimic the profile predicted by the MD simulations (Fig. \ref{simVSexpNOions_POPS}).
\todo{Tiago has to check and update this and figures in SI.}

@@

The Lipid17/Dang simulation give larger acyl chain order parameters and smaller area per molecule than the
experimental values (Fig. \ref{simVSexpNOions_POPS} and table \ref{tab:apls}),
indicating that this simulation predicts too dense POPS lipid bilayer. This is also in line with
the discrepancy observed in the X-ray form factor between this simulation and experiment (Fig. \ref{simVSexpNOions_POPS}).
In previous study \cite{NMRlipidsIV}, the Lipid17 force field with {\AA}qvist ions~\cite{aqvist90}
gave a larger area per molecule (57~\AA$^2$) for POPS being closer to the experimental value,
which is probably due to the weaker counterion binding affinity of {\AA}qvist ions.
However,  {\AA}qvist ions exhibit artificial aggregation of ions at larger salt concentrations \cite{kohagen16, chen07, NMRlipidsIV}.
Therefore, the Dang ions with more realisitic bulk properties are used here to analyze the effect of ECC to the properties of POPS lipid bilayer. 
The ECC-POPS simulation gives a good agreement with the X-ray scattering form factors and experimental
acyl chain order parameters (see Fig.~\ref{simVSexpNOions_POPS}), indicating that the bilayer dimensions and acyl chain
conformations are well described by the force field. The area per lipid from ECC-POPS simulation is slightly
smaller than the value reported from SDP model \cite{kucerka14} (table \ref{tab:apls}), but the values
agree within their error estimates.

\begin{table}[tb!] 
\centering
  \caption{Area per lipid of POPS bilayers, $A_L$, from simulations with different force fields at 298~K with \ce{Na^+} counterions. \label{tab:apls} } 
  \begin{tabular}{l|c } 
    \multicolumn{2}{c}{POPS} \\
    model          & $A_L$ / Å$^2$    \\ 
    \hline 
    Lipid17/Dang              & 53.5$\pm$ 1.7   \\ 
%    Lipid17/ff99              & 57.9$\pm$ 1.7   \\ 
%    \hline 
    ECC-POPS/ECC-ions         & 59.8$\pm$ 1.8   \\ 
%    \hline 
    experiment (SDP model) \citep{kucerka14} & 62.7$\pm$ 1.3  \\ 
%    \hline 
  \end{tabular} \\
\end{table}

Despite the success in simulating bilayer dimensions and acyl chain conformations,
lipid force fields have typically problems in capturing the correct glycerol backbone and
headgroup order parameters and conformations \cite{botan15,ollila16,NMRlipidsIV}.
This is the case also for Lipid17 POPS simulations, where the headgroup and glycerol backbone
order parameters are quite far from experimental values with both Dang (Fig. \ref{simVSexpNOions_POPS})
and {\AA}qvist \cite{NMRlipidsIV} ions. The values from ECC-POPS simulation are closer,
but not in full agreement with experiments (Fig.~\ref{simVSexpNOions_POPS}).
Especially, the experimentally observed more negative values of the $\beta$-carbon
and the larger forking (difference of order parameters between hydrogens attached to the
same carbon \cite{ollila16}) of the $\alpha$-carbon in POPS than in POPC \cite{NMRlipidsIV}
are not reproduced by neither of the models here. 
These differences are previously related
to a more rigid structure of PS headgroup and 
the stiffness of the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$ dihedral. 
Interestingely, this dihedral angle is more flexible in ECC-POPS compared to Lipid17 
suggesting that its rigidity is overestimated in Lipid17 due to the lack of electronic polarizability,
while the same dihedral parameters yield too small forking in ECC-POPS.
\todo{the dihedral to be plotted.}
Although the ECC-POPS model reproduces the lipid bilayer dimensions and
acyl chain conformations in a good agreement with experiments, 
there is a room for improvements especially in the glycerol backbone and headgroup region,
e.g. in the parameters of the associated dihedral angles. 
\cite{botan15, ollila16, Pluhackova2016, NMRlipidsIV}.
